---
title: "Group 8: Hotel Bookings"
author: "Alex Armstrong, Si Chen, Ethan Lamons"
date: "10/13/2020"
output: html_document
---
# Introduction
For this project, we used the Hotel Bookings dataset from the Tidy Tuesday R repository. Each of us took different variables and analyzed the relationships between these variables using various statistical analysis tools and graphs. The first section by Ethan Lamons compares the average daily rate (ADR) per booking with the amount and age range of the party that booked the room. A further comparison is made between the ADR and the number of kids per booking party with a linear regression analysis made for each comparison.
	For the second section, Si Chen provides data visualization in the room reservations by type. She also conducted two linear regression analysis in finding the relationship between nights stayed at hotels on weekdays or weekends and having children or babies.
	For the third section, Alex Armstrong looked at the country data. Firstly with the mean average daily rate per country, then the mean lead time per country. After that, a further comparison was made between the distribution channel and the lead time. An analysis of variance was then made between the 3 main distribution channels and their lead times.  


Loading all of the packages needed.
```{r}
install.packages("tidytuesdayR", repos = "http://cran.us.r-project.org")
library(tidytuesdayR)
library(tidyverse)
library(outliers)
library(ggdark)

install.packages("rworldmap", repos = "http://cran.us.r-project.org")
library(rworldmap)
install.packages("RColorBrewer", repos = "http://cran.us.r-project.org")
library(RColorBrewer)
```
# Part 1: Analysis Between adr and Booking Party Composition

There are two comparisons done in this section. One is between adr and the full party composition with the colors showing the breakdown of the party by age group.

The other is between adr and the number of kids. A linear regression analysis is shown at the end, and the resulting line.

## 1.1 Bringing in and modifying the dataset

First thing is to bring in the Hotel Bookings data from the Tidy Tuesday repository.
```{r}
hotels <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-02-11/hotels.csv')
hotels
```

Next thing is to modify the dataset to choose only the needed variables for the graphs and analysis. I also removed outliers and added in two new variables, "Number_of_Kids_per_Party" and "party_size". The way I determined the outliers is shown below. I also removed any adr and party_size values that equaled zero since that data is irrelevant to my analysis. 
```{r}
adr_by_party <- hotels %>%
  select(adults, children, babies, adr) %>%
  mutate(party_size = (adults + children + babies)) %>%
  mutate(Number_of_Kids_per_Party = (children + babies)) %>%
  arrange(party_size) %>%
  filter(adr < 500 & adr != 0 & party_size < 10 & party_size > 0) %>%
  na.omit(adr_by_party)
```

Here are the summary statistics of the modified dataset and two boxplots which show that adr values over 500 are outliers. The same method was applied to party_size which shows that values of party_size that are 10 or more are outliers. The black line in these boxplots show the cutoff for outliers.
```{r}
summary(adr_by_party)

ggplot(hotels) +
  aes(x = "", y = adr) +
  geom_boxplot(fill = "#000000") +
  theme_minimal() +
  geom_hline(aes(yintercept = 500), colour = "black")

hotels %>%
  filter(adr >= -50L & adr <= 1300L) %>%
  ggplot() +
  aes(x = "", y = adr) +
  geom_boxplot(fill = "#000000") +
  theme_minimal() +
  geom_hline(aes(yintercept = 500), colour = "black")
```

## 1.2 Making a graph showing relation between number of people per booking and adr

Below is a tile graph showing the relationship between the number of people per booking and adr. The colors on the graph show the number of kids per party. The white line is the result of a linear regression analysis shown in a later section. From the graph, there is a significant correlation between party size and adr with a further correlation between adr and the number of kids per party. This result lead to the next graph. A full image .png of the graph is available under the Figures folder in the repository.
```{r}
(npeople_v_adr_plot <- ggplot(adr_by_party) +
  aes(x = party_size, y = adr, colour = Number_of_Kids_per_Party) +
  geom_tile() +
  scale_color_distiller(palette = "Paired") +
  labs(x = "Number of People per Booking", y = "Average Daily Rate", title = "Number of People per Booking vs ADR Related to the Number of Kids per Party", caption = "Graph made by - Ethan Lamons") +
  dark_theme_bw() +
  theme(axis.title.x = element_text(size = 15),
        axis.title.y = element_text(size = 15, vjust = 2),
        title = element_text(size = 14),
        legend.title = element_text(size = 12)) +
  geom_abline(aes(slope = 31.64, intercept = 41.06), size = 1.5) +
  scale_x_continuous(n.break = 6) +
  annotate(geom = "segment", x = 2, y = 150, xend = 2.5, yend = 130, size = 1, colour = "black", arrow = arrow(length = unit(2, "mm"))) +
  annotate(geom = "text", x = 0.52, y = 182, label = "Linear Regression Line", hjust = "left", colour = "black") +
  annotate(geom = "text", x = 0.52, y = 167, label = "(yint = 41.06, slope = 31.64, p < 2.2e-16,", hjust = "left", colour = "black") +
  annotate(geom = "text", x = 0.52, y = 152, label = "paste(italic(R) ^ 2, \"= 0.191)\")", parse = TRUE, hjust = "left", colour = "black"))
```

## 1.3 Graph showing the relation between the number of kids per party and adr

This graph is similar to the previous except the x axis variable is the number of kids per booking, and the colors are there to make it easier to distinguish the columns. Again, the white line is from a linear regression analysis shown below. A full size .png of the graph can be found in the Figures folder.
```{r}
(nkids_v_adr <- ggplot(adr_by_party) +
  aes(x = Number_of_Kids_per_Party , y = adr, colour = Number_of_Kids_per_Party) +
  geom_tile(show.legend = FALSE) +
  scale_color_distiller(palette = "Paired") +
  labs(x = "Number of Kids per Booking", y = "Average Daily Rate", title = "Number of Kids per Booking vs ADR", caption = "Graph made by - Ethan Lamons") +
  dark_theme_bw() +
  geom_abline(slope = 40.01, intercept = 99, size = 1.5) +
  scale_x_continuous(n.break = 6) +
  theme(axis.title.x = element_text(size = 15),
        axis.title.y = element_text(size = 15, vjust = 2),
        title = element_text(size = 20)) +
  annotate(geom = "segment", x = 0.5, y = 200, xend = 1, yend = 150, size = 1, colour = "black", arrow = arrow(length = unit(2, "mm"))) +
  annotate(geom = "text", x = -.4, y = 230, label = "Linear Regression Line", hjust = "left", colour = "black") +
  annotate(geom = "text", x = -.4, y = 215, label = "(yint = 99, slope = 40.01, p < 2.2e-16,", hjust = "left", colour = "black") +
  annotate(geom = "text", x = 0.9, y = 215, label = "paste(italic(R) ^ 2, \" = 0.124)\")", parse = TRUE, hjust = "left", colour = "black"))

```

## 1.4 Linear regression analysis of the data

The linear regression lines found on each of the graphs come from the analysis shown below. First is making the linear regression variable for each comparison and then showing the summary of the linear regression values. Both graphs show a very high significance between the variables used which comes from the low p-value. The R-squared value is low, though, showing that the correlation is not very strong.
```{r}
#Corresponding to the first graph
lm.party_size <- lm(adr_by_party$adr ~ adr_by_party$party_size)

#Corresponding to the second graph
lm.nkids <- lm(adr_by_party$adr ~ adr_by_party$Number_of_Kids_per_Party)
```

```{r}
#Corresponding to the first graph
summary(lm.party_size)

#Corresponding to the second graph
summary(lm.nkids)
```

## 1.5 Creating a .txt file with the means

Finally, I created a text file that contains all the means of the variables I used. The file can be found in the Data folder.
```{r}
means_adr_by_party <- as.data.frame(lapply(adr_by_party, mean), row.names = "mean")

write.table(means_adr_by_party, file = "C:/Users/lamon/Desktop/CPSC441/Group_8/Data/adr_by_party_means.txt")
```



# PART 2: Analysis of Room Reserved and Nights stayed

## 2.1
Assigning a variable that contains the number of counts by unique room types in reserved room type and showing the counts.
```{r}
library(dplyr)
room_reservation_preference<-hotels%>% 
            count(reserved_room_type)
room_reservation_preference
```
## 2.2
Making a graph out of the value counts. From the graph, we can see that the room type A is the abusolute most reserved room type, with 85994 reservations. The type of room is not specified but I suppose room type A is the standard room.
```{r}
library(ggplot2)
room_type<- as_tibble(room_reservation_preference)
room_reservation_plot<-ggplot(room_type,mapping=aes(x=room_type$reserved_room_type, y=room_type$n))+
      geom_bar(stat = "identity")+
      labs(x="Type of room",y="Number of reservation")+
      ggtitle("Room Reservation by Type")
      
room_reservation_plot
ggsave(room_reservation_plot,device="pdf")
```

## 2.3
Conducting two linear regressions
First, we want to know whether families with children or babies stay more nights on week days
Second, we want to know whtehr families with children or babies stay more nights on weekend days
```{r}
lm.week_day_stays <- lm(hotels$stays_in_week_nights ~ hotels$children+hotels$babies)
lm.weekend_day_stays <- lm(hotels$stays_in_weekend_nights ~ hotels$children+ hotels$babies)
```
Outputting summaries for the two regressions
We can see from the regression results that both children and babies are very significant in explaining the dependent variables that are weekday stays and weekend stays. However, both adjusted R-squared are very low. This suggests that whether having children or babies can not explain much of the variance in nights staying on weekdays or weekends.
```{r}
summary(lm.week_day_stays )
summary(lm.weekend_day_stays )
```

## 2.4
Saving the means of variables I used to a text file
```{r}
means_hotel_stays <- hotels%>%
   select(reserved_room_type,stays_in_week_nights,stays_in_weekend_nights,children,babies)%>%
   na.omit(means_hotel_stays)

means_hotel_stays <- as.data.frame(lapply(means_hotel_stays, mean), row.names = "mean")


write.table(means_hotel_stays, file = "hotel_stays_means.txt")
```


# PART 3: Countries / Lead Time - Alex:

I was initially interested in the country by country data, which is difficult to visualize with a simple graph as there are hundreds of countries in the data-set, so I used the 'rworldmap' package. 
The hotels data set includes the country data in the form of iso3 country codes, which allows the hotel data to be joined to the correct country on the world map. 
Visualizing the data in map form allows us to see that the average daily rate (defined by dividing the sum of all lodging transactions by the total number of staying nights within a country) appears to be higher in the U.S. and other countries such as Kazakhstan.
First, I had to manipulate the data by using group_by to group the data for each country. Then I took the mean of the average daily rate for each country.

```{r}
by_country <- hotels %>% 
  group_by(country) %>% 
  summarise(mean_adr=mean(adr))
print(by_country)
mapped_data <- joinCountryData2Map(by_country, 
        joinCode = "ISO3", 
        nameJoinColumn = "country")

par(mai=c(0,0,0.2,0), xaxs = "i", yaxs = "i")

mapParams <- mapCountryData(mapped_data, 
        nameColumnToPlot = "mean_adr",
        mapTitle = "Mean Average Daily Rate by Country",
        addLegend=F)

do.call(addMapLegend,
        c(mapParams,
          legendLabels='limits',
          legendWidth=1.0,
          legendIntervals='data'
          ))
```

Using the same tools, I started looking into the Lead Time data, which states how far in advance bookings are made. Once again I used the 'rworldmap' package. 
This time I made a legend box, which shows the categories of the data much better.  
I manually created the bounds for the legend.
I also used a different colour scheme using the RColorBrewer package.


```{r}
by_country2 <- hotels %>% 
  group_by(country) %>%
  summarise(mean_leadtime=mean(lead_time))

colourPalette2 <- brewer.pal(n=7,name='RdPu')

mapped_data2 <- joinCountryData2Map(by_country2, 
        joinCode = "ISO3", 
        nameJoinColumn = "country")

par(mai=c(0,0,0.2,0), xaxs = "i", yaxs = "i")

mapParams2 <- mapCountryData(mapped_data2, 
        nameColumnToPlot = "mean_leadtime",
        mapTitle = "Lead Time by Country",
        colourPalette = colourPalette2,
        addLegend = F)

mapParams2$legendText <-
  c('0 - 22.72',
    '22.72-42.89',
    '42.89-57.85',
    '57.85-77.21',
    '77.21- 94.85',
    '94.85-123.96',
    '123.96+')

do.call(addMapLegendBoxes,
        c(mapParams2,
          title='Lead Times (Days)'))
```

Further exploring the lead time variable showed there appeared to be a difference in lead time depending on the distribution method. 
To measure this, I created a bar plot. This time it shows the average lead time within different distribution channels.
First I grouped by distribution channel again and then caluclated the mean and standard devaitions for each distribution channel.

```{r}
by_distribution_channel<-hotels %>% 
  group_by(distribution_channel) %>%
  summarise(mean_dist_leadtime=mean(lead_time),
            sd_dist_leadtime=sd(lead_time))

ggplot(by_distribution_channel, aes(x=distribution_channel, y=mean_dist_leadtime)) + 
  geom_bar(stat='identity', aes(fill=distribution_channel), col='#000000', na.rm=T) + 
  ylim(c(0, 250)) + 
  labs(y="Lead Time (Days) +/- SD", 
       x="Distribution Type", 
       title="Average Lead Time per Distribution Channel" 
       )+
  geom_errorbar(aes(ymin=mean_dist_leadtime,  # I left the ymin as the mean, as most of the ymin values were below zero, so I wanted to exclude them from the graph
                    ymax=mean_dist_leadtime + sd_dist_leadtime),
                    width=.8,
                    position='dodge') +
  theme(plot.title=element_text(hjust = 0.5), legend.position = 'none')

```

For the statistical analysis, I was keen to see whether there was any significant difference between the 3 main distribution channels shown in the graph above.
To do this, I first isolated the 3 most popular distribution channels (direct, coporate & TA/TO).
Then I carried out an analysis of variance across the 3 different channels.
The results show there is significance.

```{r}
distribution_channel_analysis<-hotels %>% 
  select(distribution_channel, lead_time) %>%
  mutate(distribution_channel = factor(distribution_channel,
                                       levels = c("Direct", "Corporate", "TA/TO")))

# Compute the analysis of variance
res.aov <- aov(distribution_channel_analysis$lead_time ~ distribution_channel_analysis$distribution_channel)
# Summary of the analysis
summary(res.aov)
```

Finally, I saved the mean variables to a file. 
I used the 'by_country', 'by_country2' and 'by_distribution_channel' data frames from earlier, as they already list the means.
by_country displays the mean average daily rate for each country
by_country2 displays the mean lead_time for each country
by_distribution_channel displays the mean lead_time for each distribution channel

```{r}
#write.table(by_country, file = "C:\Users\AlexA\OneDrive\Documents\R\Group_8\Data\Alex_Means.txt")
#write.table(by_country2, file = "C:\Users\AlexA\OneDrive\Documents\R\Group_8\Data\Alex_Means.txt", append=T)
#write.table(by_distribution_channel, file = "C:\Users\AlexA\OneDrive\Documents\R\Group_8\Data\Alex_Means.txt", append=T)
#>>>>>>> 861545224ed1f38a83eeb39a680cd198bc55026a
